<!--
  %\VignetteIndexEntry{More igraph visualization with Lucy}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
-->

# Useful visualizations using lucy

```{r, message=FALSE}
#devtools::install_github("robertness/lucy")
library(lucy)
set.seed(20)
```

## Graphviz + igraph

The [*Rgraphviz*](http://www.bioconductor.org/packages/release/bioc/html/Rgraphviz.html) package makes very useful layouts for graphs.  A basic Rgraphviz visualization is implemented in the *igraphviz* function.

```{r}
igraphviz(ba.game(10))
```

Rgraphviz is especially useful in the case when there is a hierarchy between graph vertices that can be illustrated with a top-down ordering in the layout.   For example, lucy's *layerDAGs* function stacks layers of highly-connected directed acyclic graphs on top of one another. 

```{r}
g <- layer_DAGs(6, 3)
igraphviz(g)
```

The *plot_path* function takes advantage of this hierarchy visualization technique by coloring all edges lying on paths from an upstream vertex to a downstream vertex.  You can clearly see flow along the paths from top to bottom.

```{r}
plot_path(g, 6, 13)
```
